The Quench-Action method in integrable spin chains: A Monte Carlo approach

### 2015-09-10 draft/fs_qa_v1.tex
revised 10/09/2015

\documentclass[11pt]{iopart}
\usepackage[english]{babel}
\pdfoutput=1
%\usepackage{amsmath}
\usepackage{wasysym}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{sidecap}
\usepackage{bm}% bold math
\usepackage{tikz}

@@

\author{Vincenzo Alba$^1$, Pasquale Calabrese$^1$}
\address{$^1$ International School for Advanced Studies (SISSA),
Via Bonomea 265, 34136, Trieste, Italy,
INFN, Sezione di Trieste}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bethe-Takahashi equations}

@@

@@

Here we restrict to the parity-invariant eigenstates. These are the only 
eigenstates with non-zero overlap with the Neel state. Parity-invariant 
eigenstates contain only pairs of rapidities with opposite sign.

@@

The overlap between the generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state is obtained fro the overlap with the Neel state as 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\lambda_j\}_{j=1}^M\rangle=\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)\langle N|\{\lambda_j\}_{j=1}^M\rangle
\end{equation}
%
The mutliplicative factor in~\eref{ov-MG}, using the string hypothesis for the generic 
$n$-string is rewritten as 
%
\begin{equation}
\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)=2^n\prod\limits_{k=0}^{\lfloor 
n/2\rfloor}\frac{1}{[\lambda_{n;\gamma}^2+(2k+(1-(-1)^n)/2)^2]^2}
\end{equation}
%

@@

@@

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel}
\end{center}
\caption{ Overlap sum rules for the Neel state. (a) The overlap sum rule 
 $\sum_{\lambda}|\langle\lambda|N\rangle|^2=1$, with $|N\rangle$ 
 the Neel state and $|\lambda\rangle$ the eigenstates  of the $XXX$ 
 spin chain. The $x$-axis shows $1/L$, with $L$ the chain length. The 
 circles are Bethe ansatz results for chains up to $L=36$. The results 
 are obtained via a full scanning of the chain Hilbert space. Only the 
 eigenstates with no zero-momentum strings are considered. The dash-dotted 
 line is a fit to $A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Inset: 
 The same data as in the main Figure plotted versus $1/L^{1/2}$. (b) 
 The same as in (a) for the sum rule $\sum_{\lambda}|\langle
 \lambda|N\rangle|^2e(\lambda)=1/2$, with $e(\lambda)$  the 
 eigenstates energy density. 
}
\label{fig2-neel-sr}
\end{figure}
%##################################################################

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo implementation of the Quench-Action approach}

For the finite-size Heisenberg spin chain the Quench-Action expectation 
values can be obtained by sampling the eigenstates of the chain using 
Monte Carlo. One starts with an eigenstate of the $XXX$ chain with 
$N_{\infty}$ infinite rapidities and $M\equiv L/2-N_\infty$ finite ones. 
The state is identified by a parity-invariant Bethe quantum number 
configuration ${\mathcal C}$ and by the corresponding parity-invariant 
rapidities $\{\lambda\}$ as $|\lambda\rangle$. The string content 
associated with the finite rapidities is denoted as ${\mathcal S}$. The Monte 
Carlo procedure consists of four steps as follows: 
%
\begin{enumerate}
\item[\circled{1}] Choose a new number of finite rapidities $M'$ with 
probability
%
\begin{equation}
\label{PM}
{\mathcal P}(M')=\frac{\widetilde Z'_{Neel}(L,M')}{\widetilde{Z}_{Neel}(L)}. 
\end{equation}
%
\item[\circled{2}] Choose a new a new string content ${\mathcal S}'\equiv
\{s_1',\dots,s_{M'}'\}$ with probability ${\mathcal P}'(M',{\mathcal S}')$
%
\begin{equation}
\label{PS}
{\mathcal P}'(M',{\mathcal S}')=\frac{1}{\widetilde Z'_{Neel}(L,M')}
\prod_{n=1}^{M'}B\Big(\frac{L}{2}-\frac{1}{2}\sum\limits_{m=1}^{M'}t_{nm}
s'_m,s'_n\Big).
\end{equation}
%
\item[\circled{3}] Generate a new parity-invariant quantum number configuration 
${\mathcal C}'$ compatible with the ${\mathcal  S}'$ obtained in step $\circled{2}$. 
Solve the corresponding BGT equations~\eref{bt_eq}, finding a new eigenstate 
$|\lambda'\rangle$. 
\item[\circled{4}] Calculate the overlap $\langle\lambda'|N\rangle$ between the 
new eigenstate and the N\'eel state and accept the eigenstate with the Metropolis 
probability 
%
\begin{equation}
\label{metropolis}
{\mathcal P}''_{\lambda\to\lambda'}=\textrm{Min}\Big\{1,\exp\Big(-
2\Re({\mathcal E}'-{\mathcal E})\Big)\Big\}, 
\end{equation}
%
where ${\mathcal E}'\equiv-\log\langle\lambda'|N\rangle$, and similarly for 
${\mathcal E}$. 
\end{enumerate}
%

@@

The validity of the Monte Carlo approach for simulating the Quench-Action results 
is demonstrated in Figure~\ref{fig5-neel-sum-rules}. The Figure focuses on the 
N\'eel sum rules for the conserved charges $Q_2$ and $Q_4$. Specifically, 
panel (a) shows the sum rule $\langle Q_2\rangle/L=1/2$ (energy conservation). 
%
\begin{equation}
\langle N|\hat Q_n|N\rangle=\sum\limits_{\lambda}|\langle N|\lambda\rangle|^2
Q_n
\end{equation}
%

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/QAMC_Obs_Neel}
\end{center}
\caption{The overlap sum rules for the Neel state: Numerical results 
 obtained using the Hilbert space Monte Carlo sampling approach. (a) 
 The energy sum rule $\langle N|Q_2|N\rangle/L=-1/2$, with $Q_2/L$ 
 the Hamiltonian density. We plot $\sum_{\lambda}|\langle\lambda|
 N\rangle|^2Q_2(\lambda)/L=1/2$, with $|\lambda\rangle$ the 
 eigenstates of the $XXX$ chain, versus the inverse chain length $1/L$. 
 The symbols are Monte Carlo data obtained by sampling the eigenstates 
 of the $XXX$ chain. The dash-dotted line is the expected result. The 
 dashed line is a fit to the behavior $-1/2+A/L+B/L^2$, with $A,B$ 
 fitting parameters. (b) The energy fluctuations sum rule $\sigma^2(Q_2)/
 L\equiv(\langle N|Q_2^2|N\rangle-\langle N|Q_2|N\rangle^2)/L=1/4$. The 
 horizontal line is the expected result. (c)(d) Same as in (a)(b) for 
 the charge $Q_4$ and its fluctuations. 
}
\label{fig5-neel-sum-rules}
\end{figure}
%##################################################################

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.8\textwidth]{./draft_figs/Neel_rho}
\end{center}
\caption{ The steady-state root densities $\rho_1(x)$ (panel (a)) and 
 $\rho_2(x)$ (panel (b)), plotted versus the root $x$. The histograms 
 are Monte Carlo data for the Heisenberg chain with $L=56$ sites and 
 $\sim 10^7$ Monte Carlo steps. The $y$-axis is divided by $10^6$ for 
 convenience. The continuous lines are the analytical Quench-Action 
 results in the thermodynamic limit. 
}
\label{fig6-neel-roots}
\end{figure}
%##################################################################

@@

>>>>

\documentclass[11pt]{iopart}
\usepackage[english]{babel}
\pdfoutput=1
%\usepackage{amsmath}
\usepackage{wasysym}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{sidecap}
\usepackage{bm}% bold math
\usepackage{tikz}
%\usepackage[backend=bibtex]{biblatex}

@@

\author{Authors}
%\address{$^1$ International School for Advanced Studies (SISSA),
%Via Bonomea 265, 34136, Trieste, Italy,
%INFN, Sezione di Trieste}

@@

In the thermodynamic limit it is natural to assume that the steady-state expectation 
values are obtained using the so-called diagonal ensemble, which is defined as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle=\sum\limits_{\alpha}|\langle\Psi_0|\alpha\rangle|^2
\langle\alpha|{\mathcal O}|\alpha\rangle. 
\end{equation}
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bethe-Gaudin-Takahashi (BGT) equations}

@@

In the thermodynamic limit the BGT equations become a set of coupled integral 
equations given as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'). 
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The conserved charges}

The $XXX$ chain exhibits an extensive number of local conserved charges $Q_n$. 
These are defined as 
%
\begin{equation}
\label{Q-def}
\left.Q_{j+1}\equiv\frac{i}{(j-1)!}\frac{d^j}{dy^j}\log\tau
(y)\right|_{y=i},
\end{equation}
%
where $y$ is a spectral parameter and $\tau(y)$ is the eigenvalue of the 
so-called transfer matrix in the Algebraic Bethe Ansatz framework. The 
analytic expression of $\tau(y)$ for a given eigenstate identified by 
a set of rapidities $\{\lambda_\alpha\}$ is given as 
%
\begin{equation}
\label{tau}
\tau(y)\equiv\Big(\frac{y+i}{2}\Big)^L\prod\limits_\alpha\frac{y-\lambda_\alpha-2i}
{y-\lambda_\alpha}+\Big(\frac{y-i}{2}\Big)^L\prod\limits_\alpha\frac{y-\lambda_\alpha
+2i}{y-\lambda_\alpha}.
\end{equation}
%
Using the string hypothesis and~\eref{Q-def} and~\eref{tau}, one obtains tha for 
a generic $n$-string with string center $\lambda$ the first few charges have the 
simple expressions
%
\begin{eqnarray}
Q_2=-\frac{2n}{\lambda^2+n^2}\\
Q_3=-\frac{4n\lambda}{(\lambda^2+n^2)^2}\\
Q_4=\frac{2n(n^2-3\lambda^2)}{(n^2+\lambda^2)^3}\\
Q_5=\frac{8n\lambda(n^2-\lambda^2)}{(n^2+\lambda^2)^4}\\
Q_6=-\frac{2n(5\lambda^4-10n^2\lambda^2+n^4)}{(n^2+\lambda^2)^5}
\end{eqnarray}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{N\'eel expectation values}
\label{Q-Neel}

The expectation values of the $XXX$ chain conserved charges over the N\'eel 
state are given as

@@

Here we focus on the overlap with the zero-momentum N\'eel state $|N\rangle$, 
defined as 
%
\begin{equation}
\label{neel}
|N\rangle\equiv\frac{1}{\sqrt{2}}\Big(|\uparrow\downarrow\rangle^{\otimes L/2}
+|\downarrow\uparrow\rangle^{\otimes L/2}\Big).
\end{equation}
%
Here we restrict to the parity-invariant eigenstates. These are the only 
eigenstates with non-zero overlap with the Neel state. Parity-invariant 
eigenstates contain only pairs of rapidities with opposite sign.

@@

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}
\end{equation}
%
The overlap between the generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state is obtained fro the overlap with the Neel state as 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\lambda_j\}_{j=1}^M\rangle=\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)\langle N|\{\lambda_j\}_{j=1}^M\rangle
\end{equation}
%
The mutliplicative factor in~\eref{ov-MG}, using the string hypothesis for the generic 
$n$-string is rewritten as 
%
\begin{equation}
\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)=2^n\prod\limits_{k=0}^{\lfloor 
n/2\rfloor}\frac{1}{[\lambda_{n;\gamma}^2+(2k+(1-(-1)^n)/2)^2]^2}
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quench Action method}
\label{quench-action}

For integrable models the Quench-Action approach allows to extract exactly both 
the steady state after the quench and the relaxation dynamics to the steady state.

In the thermodynamic limit the sum over the model eigenstates is recast into 
a functional integral over the so-called root distributions $\pmb{\rho}\equiv
\{\rho_n(\lambda)\}_{n=1}^\infty$ as
%
\begin{equation}
\label{eig-sum}
\sum\limits_{\alpha}\rightarrow\int{\mathcal D}\pmb{\rho}e^{S_{YY}(\pmb{\rho})}. 
\end{equation}
%
Here ${\mathcal D}\pmb{\rho}\equiv\prod_{n=1}^\infty{\mathcal D}\rho_n(\lambda)$, 
$\rho_n(\lambda)$ denote the distribution of the $n$-strings BGT roots, and 
$S_{YY}(\pmb{\rho})$ is the Yang-Yang entropy, which counts the number of 
Bethe states leading to the same $\pmb{\rho}$ in the thermodynamic limit. 
Formally, $\rho_n(\lambda)$  are defined as $\rho_n(\lambda)\equiv\lim_{L\to
\infty}(\lambda_{n;\gamma+1}-\lambda_{n;\gamma})^{-1}$.

Using~\eref{eig-sum} the diagonal ensemble~\eref{d-ensemble} becomes 
%
\begin{equation}
\label{qa-d-ensemble}
\fl
\quad\langle{\mathcal O}\rangle=\int{\mathcal D}\pmb{\rho}(\lambda)
\exp\Big(2\Re\log\langle\Psi_0|\pmb{\rho}(\lambda)\rangle 
+S_{YY}(\pmb{\rho})\Big)\langle\pmb{\rho}(\lambda)|{\mathcal O}|\pmb{\rho}
(\lambda)\rangle
\end{equation}
%
Note that in~\eref{qa-d-ensemble} it is assumed that in the thermodynamic limit 
the eigenstate expectation values $\langle\alpha|{\mathcal O}|\alpha\rangle$ 
become smooth functionals  $\langle\pmb{\rho}(\lambda)|{\mathcal O}|\pmb{\rho}
(\lambda)\rangle$ of the root distributions $\pmb{\rho}$.

Similarly, in~\eref{qa-d-ensemble} $\langle\Psi_0|\pmb{\rho}(\lambda)\rangle$ 
denotes the Bethe states overlap density with the prequench state $\Psi_0$.

The strategy is to evaluate the functional integral in~\eref{qa-d-ensemble} 
using the saddle point approximation.

One has then to minimize the functional ${\mathcal S}(\pmb{\rho})$ defined 
as 
%
\begin{equation}
L{\mathcal S}(\pmb{\rho})\equiv 2\Re\log\langle\pmb{\rho}|\Psi_0
\rangle+S_{YY}(\pmb{\rho}(\lambda)). 
\end{equation}
% 
under the constraint that the BGT equations hold. For the quench with initial 
state the N\'eel state the resulting saddle point root distributions 
$\rho_n^*(\lambda)$ can be obtained analytically. The first few root distributions 
are given as 
%
\begin{eqnarray}
\label{rho_sp}
\fl\quad\rho^*_1(\lambda)=\frac{8(4+\lambda^2)}{\pi(19+3\lambda^2)(1+6\lambda^2+
\lambda^4)}\\
\fl\quad\rho_2^*(\lambda)=\frac{8\lambda^2(9+\lambda^2)(4+3\lambda^2)}{\pi(2+\lambda^2)
(16+14\lambda^2+\lambda^4)(256+132\lambda^2+9\lambda^4)}\\
\fl\quad\rho_3^*=\frac{8(1+\lambda^2)^2(5+\lambda^2)(16+\lambda^2)(21+\lambda^2)}
{\pi(19+3\lambda^2)(9+624\lambda^2+262\lambda^4+32\lambda^6+\lambda^8)
(509+5\lambda^2(26+\lambda^2))}.
\end{eqnarray}
%

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel_overlaps}
\end{center}
\caption{ 
}
\label{fig0}
\end{figure}
%##################################################################

@@

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/Neel}
\end{center}
\caption{ Overlap sum rules for the Neel state: The role of the zero-momentum 
 strings. (a) The overlap sum rule $\sum_{\lambda}|\langle\lambda|N\rangle|^2=1$, 
 with $|N\rangle$ the Neel state and $|\lambda\rangle$ the eigenstates  of 
 the $XXX$ spin chain. The $x$-axis shows $1/L$, with $L$ the chain length. 
 The circles are Bethe ansatz results for chains up to $L=36$. The results 
 are obtained via a full scanning of the chain Hilbert space. Only the 
 eigenstates with no zero-momentum strings are considered. The dash-dotted 
 line is a fit to $A/L^{1/2}+B/L$, with $A,B$ fitting parameters. Inset: 
 The same data as in the main Figure plotted versus $1/L^{1/2}$. (b) 
 The same as in (a) for the sum rule $\sum_{\lambda}|\langle
 \lambda|N\rangle|^2e(\lambda)=1/2$, with $e(\lambda)$  the 
 eigenstates energy density. 
}
\label{fig2-neel-sr}
\end{figure}
%##################################################################

@@

For the N\'eel state, for any finite $L$ Eq.~\eref{sum} gives $Q_0=1,-1/2$ for 
$\hat{Q}=\mathbb{I}$ and $\hat Q={\mathcal H}$, respectively.

Figure~\ref{fig2-neel-sr} plots the numerical results obtained via a full 
scanning of the Hilbert space of the Heisenberg chain.

Precisely, the numerical data are obtained by generating all the possible string 
contents ${\mathcal S}$ compatible with the parity-invariance constraint

Panel (a) and (b) in the 
Figure plots the sum rule~\eref{sum} for $\hat Q=\mathbb{I}$ and $\hat Q={\mathcal 
H}$, respectively. The circles are obtained considering all the eigenstates of the 
Heisenberg spin chain with $L$ sites, excluding eigenstates containing 
zero-momentum strings.

The total number of eigenstates $\widetilde Z_{Neel}$ containing no zero-momentum 
strings is given as 
%
\begin{equation}
\label{ztilde}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big).
\end{equation}
%
The proof of\eref{ztilde} is reported in~\ref{app-2}.

@@

The largest chain size considered is $L=36$. The data are plotted versus $1/L$.

Interestingly, both the sum rules (panels (a)(b)) are not saturated, as expected 
since the eigenstates corresponding to zero-momentum strings are excluded in the 
sum in the right-hand-side in~\eref{sum}.

Interestingly, both sum rules exhibit vanishing behavior upon increasing the 
chain size. The dash-dotted lines in the Figures are fits to the behavior 
$A/L^{1/2}+B/L$, with $A,B$ fitting parameters, and perfectly describe the 
behavior of the data.

The vanishing behavior as $\propto 1/L^{1/2}$ of the sum rules reflects the vanishing 
of the fraction of non-zero momentum string eigenstates in the thermodynamic limit ad 
%
\begin{equation}
\label{beh}
\frac{\widetilde Z_{Neel}}{Z_{Neel}}\propto\frac{4}{\sqrt{\pi L}}, 
\end{equation}
% 
with $Z_{Neel}$ being the total number of eigenstates corresponding to parity 
invariant BGT quantum numbers configurations.

One should observe that the asymptotic, i.e., at large $L$, behavior~\eref{beh} 
is not generic, meaning that different initial states $|\Psi_0\rangle$ might 
give different behaviors.

This is illustrated in Figure~\ref{fig3-dimer-sr}, focusing on the Majumdar-Ghosh 
(MG) state. Panels (a) and (b) plot the sum rules with $\hat Q=\mathbb{I}$ and 
$\hat Q={\mathcal H}$ (same as in Figure~\ref{fig2-neel-sr}). As for the N\'eel 
state only eigenstates containing no zero-momentum strings are consdidered. 
Note that their number $\widetilde Z_{MG}$ (see~\ref{app-2}) is now given as 
%
\begin{equation}
\widetilde Z_{MG}=B\Big(\frac{L}{2},\frac{L}{4}\Big)-B\Big(\frac{L}{2},
\frac{L}{4}-1\Big). 
\end{equation}
%
Similar to Figure~\ref{fig2-neel-sr}, due to the exclusion of the zero-momentum 
strings, the saturation rules exhibit vanishing behavior in the thermodynamic 
limit. However, in contrast with the N\'eel case (see Figure~\ref{fig2-neel-sr}), 
one has the behavior $Q_0\sim 1/L$, as confirmed by the fits (dash-dotted lines 
in Figure~\ref{fig3-dimer-sr}).

Similar to the N\'eel case this reflects the same vanishing behavior of the 
eigenstates containing no zero-momentum strings $\widetilde Z_{MG}/Z_{MG}$ as 
%
\begin{equation}
\frac{\widetilde Z_{MG}}{Z_{MG}}=\frac{4}{4+L}. 
\end{equation}
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo implementation of the Quench-Action approach}
\label{mcqa-sec}

For the finite-size Heisenberg spin chain the Quench-Action expectation 
values can be obtained by sampling the eigenstates of the chain using 
Monte Carlo. One starts with an eigenstate of the $XXX$ chain with 
$N_{\infty}$ infinite rapidities and $M\equiv L/2-N_\infty$ finite ones. 
The state is identified by a parity-invariant Bethe quantum number 
configuration ${\mathcal C}$ and by the corresponding parity-invariant 
rapidities $\{\lambda\}$ as $|\lambda\rangle$. The string content 
associated with the finite rapidities is denoted as ${\mathcal S}$. The Monte 
Carlo procedure consists of four steps as follows: 
%
\begin{enumerate}
\item[\circled{1}] Choose a new number of finite rapidities $M'$ with 
probability
%
\begin{equation}
\label{PM}
{\mathcal P}(M')=\frac{\widetilde Z'_{Neel}(L,M')}{\widetilde{Z}_{Neel}(L)}. 
\end{equation}
%
\item[\circled{2}] Choose a new a new string content ${\mathcal S}'\equiv
\{s_1',\dots,s_{M'}'\}$ with probability ${\mathcal P}'(M',{\mathcal S}')$
%
\begin{equation}
\label{PS}
{\mathcal P}'(M',{\mathcal S}')=\frac{1}{\widetilde Z'_{Neel}(L,M')}
\prod_{n=1}^{M'}B\Big(\frac{L}{2}-\frac{1}{2}\sum\limits_{m=1}^{M'}t_{nm}
s'_m,s'_n\Big).
\end{equation}
%
\item[\circled{3}] Generate a new parity-invariant quantum number configuration 
${\mathcal C}'$ compatible with the ${\mathcal  S}'$ obtained in step $\circled{2}$. 
Solve the corresponding BGT equations~\eref{bt_eq}, finding a new eigenstate 
$|\lambda'\rangle$. 
\item[\circled{4}] Calculate the overlap $\langle\lambda'|N\rangle$ between the 
new eigenstate and the N\'eel state and accept the eigenstate with the Metropolis 
probability 
%
\begin{equation}
\label{metropolis}
{\mathcal P}''_{\lambda\to\lambda'}=\textrm{Min}\Big\{1,\exp\Big(-
2\Re({\mathcal E}'-{\mathcal E})\Big)\Big\}, 
\end{equation}
%
where ${\mathcal E}'\equiv-\log\langle\lambda'|N\rangle$, similarly for 
${\mathcal E}$, and $\Re$ denotes the real part. 
\end{enumerate}
%

Afte some necessary thermalization steps the sequence $1-4$ generates eigenstates 
of the $XXX$ chain distributed according to the Quench-Action probability 
distribution.

For a generic observable ${\mathcal O}$, its Quench-Action expectation value 
$\langle{\mathcal O}\rangle_{QA}$ is obtained as the arithmetic average of the 
expectation value over the eigenstates sampled in the Monte Carlo as 
%
\begin{equation}
\langle{\mathcal O}\rangle_{QA}=\frac{1}{N_{mcs}}\sum\limits_{\lambda}
\langle\lambda|{\mathcal O}|\lambda\rangle, 
\end{equation}
%
with $N_{mcs}$ being the total number of Monte Carlo steps.

@@

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\textwidth]{./draft_figs/QAMC_Obs_Neel}
\end{center}
\caption{The overlap sum rules for the Neel state: Numerical results 
 obtained using the Hilbert space Monte Carlo sampling approach. (a) 
 The energy sum rule $\langle N|Q_2|N\rangle/L=-1/2$, with $Q_2/L$ 
 the Hamiltonian density. We plot $\sum_{\lambda}|\langle\lambda|
 N\rangle|^2Q_2(\lambda)/L=1/2$, with $|\lambda\rangle$ the 
 eigenstates of the $XXX$ chain, versus the inverse chain length $1/L$. 
 The symbols are Monte Carlo data obtained by sampling the eigenstates 
 of the $XXX$ chain. The dash-dotted line is the expected result. The 
 dashed line is a fit to the behavior $-1/2+A/L+B/L^2$, with $A,B$ 
 fitting parameters. (b) The energy fluctuations sum rule $\sigma^2(Q_2)/
 L\equiv(\langle N|Q_2^2|N\rangle-\langle N|Q_2|N\rangle^2)/L=1/4$. The 
 horizontal line is the expected result. (c)(d) Same as in (a)(b) for 
 the charge $Q_4$ and its fluctuations. 
}
\label{fig5-neel-sr}
\end{figure}
%##################################################################

In this section we numerically demonstrate that the only effect of neglecting 
the zero-momentum strings are finite-size corrections.

The validity of the Monte Carlo approach for simulating the Quench-Action 
results is demonstrated in Figure~\ref{fig5-neel-sr}. The Figure focuses on 
the N\'eel sum rules for the conserved charges $Q_2$ and $Q_4$. Specifically, 
panel (a) shows the sum rules for $\hat Q=\hat Q_2$, $\hat Q=\hat Q_4$, 
and the corresponding fluctuations $\sigma^2(Q_n)\equiv \langle N|\hat Q_n^2|
N\rangle-\langle N|\hat Q_n|N\rangle^2$. 
%
\begin{equation}
\langle N|\hat Q_n|N\rangle=\sum\limits_{\lambda}|\langle N|\lambda\rangle|^2
Q_n.
\end{equation}
%
Note that the N\'eel sum rule for the identity $\mathbb{I}$ is trivially satisfied, 
and it corresponds to the normalization of the Monte Carlo probability. 
Figure~\ref{fig5-neel-sr} (a) plots the N\'eel sum rule for the energy density $Q_2/L$. 
The circles are Monte Carlo data obtained using the procedure outlined in~\ref{mcqa-sec} 
for several Heisenberg spin chains. The largest chain simulated is for $L=56$ sites. 
The data correspond to a simulation with $\sim 10^7$ Monte Carlo step (mcs). 
The expected result $\langle N|\hat Q_2|N\rangle/L=-1/2$ is shown as dash-dotted 
line.

In stark contrast with Figure~\ref{fig2-neel-sr} and Figure~\ref{fig3-dimer-sr}, 
while for small chains the sum rule is violated the data strongly suggest that 
in the thermodynamic limit the sum rule saturates the expected theoretical 
value.

Furthermore, the data suggest the behavior $\propto 1/L$, as confirmed by the fit 
to $-1/2+A/L+b/L^2$ (dashed line in the Figure), with $A,B$ fitting parameters.

The same qualitative behavior is demonstrated in panel (b) for the energy 
fluctuations. Interestingly, for $L=56$ the Monte Carlo result is compatible 
within the error bar with the expected asymptotic one.

Finally, panels (c)(d) show the same results as for (a)(b) for the conserved 
charge $Q_4$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Quench-Action root distributions}
\label{MCQA-root}

%##################################################################
\begin{figure}[t]
\begin{center}
\includegraphics[width=.95\textwidth]{./draft_figs/Neel_rho}
\end{center}
\caption{ The Quench-Action steady state root ddistributions $\rho_1(x)$ 
 and $\rho_2(x)$: Monte Carlos results. (a) The histograms of the $1$-string 
 rapidities sampled in the Monte Carlo. The $x$-axis shows the $1$-string BGT 
 root $\lambda$. The data are for a chain with $L=56$ sites and $N_{mcs}\sim 
 10^7$ Monte Carlo steps. The data are divided by a factor $10^6$ for plotting 
 convenience. The width of the histogram bin is $\Delta\lambda\sim 0.07$. (b) 
 The same as in (a) for the $2$-string roots sampled in the Monte Carlo. (b) 
 The extracted $1$-string root distribution $\rho_1(\lambda)$ plotted versus 
 $\lambda$ for two chains with $L=48$ and $L=56$ (diamond and circles, respectively). 
 The full line is the Quench-Action analytic result in the thermodynamic limit. (d) 
 The same as in (c) for the $2$-string root density $\rho_2(\lambda)$. In both 
 (c)(d) the oscillations are finite-size effects, whereas the error bars 
 show the statistical Monte Carlo error. 
}
\label{fig6-neel-roots}
\end{figure}
%##################################################################

The BGT root distributions $\rho_n(\lambda)$ identifying the Quench-Action saddle 
point can be extracted from the Monte Carlo simulation as shown in~\cite{alba-2015}.

Figure~\ref{fig6-neel-roots} displays the first two root distributions 
$\rho_1(\lambda)$ (panel (a)) and $\rho_2(\lambda)$ (panel (b)), plotted agains 
the BGT root $\lambda$. The data are the histograms of the BGT roots sampled 
in the Monte Carlo. The data are for a chain with $L=56$ sites and 
$N_{mcs}\sim 10^7$ Monte Carlo steps. All the data are divided by a factor 
$10^6$ for plotting convenience. The width of the histogram bins $\Delta\lambda$ 
is $\Delta\lambda\approx0.02$ and $\Delta\lambda\approx0.001$ for $\rho_1(\lambda)$ 
and $\rho_2(\lambda)$, respectively.

The histogram fluctuations are due to $N_{mcs}$ being finite, and are expected 
to vanish in the limit $N_{mcs}\to\infty$.

The continuous lines is the expected analytic result in the thermodynamic limit.

Clearly, $\rho_1(\lambda)$ is in good agreement with the Monte Carlo data in the 
whole range $-2\le\lambda\le2$ considered. 
On the other hand, larger deviations from the theoretical result are present 
for $\rho_2(\lambda)$. These deviations are larger on the tails of the 
distribution. This is expected since large $\lambda$ correspond to large 
quasimomenta, which are more sensitive to the finite size of the chain.

@@

Note that the total number $Z_{MG}$ of parity-invariant eigenstates having non 
zero overlap with the Majumdar-Ghosh state is obtained from Eq~\eref{N-count-p} 
replacing $\ell=L/2$, to obtain 
%
\begin{equation}
\label{p-inv-mg}
Z_{MG}=B\Big(\frac{L}{2}-1,\frac{L}{4}-1\Big)+B\Big(\frac{L}{2}-1,\frac{L}{4}
\Big). 
\end{equation}
%
Physically, this is due to the fact that the 
Majumdar-Ghosh state is invariant under $SU(2)$ rotations, which implies 
that only eigenstates with zero total spin $S=0$ can have non zero overlap.
